Busch, Khakhalin. Intrinsic temporal tuning

### 2018-12-03 reviews and responses.md

Correction accepted.

> 4.	The statistical tests applied throughout the paper are unclear and need more explanation. The number of degrees of freedom (e.g, 667) seems far too high, suggesting that multiple recordings from the same cell under different conditions are being treated as independent N. However recordings from the same cell under different temporal or amplitude conditions are far from independent measurements. It seems that a two-way or even three-way ANOVA might be the best approach for examining the variables of conditioning stimulus and response property.

The number of degrees of freedom in question comes from a full linear model, and does mean that we treated different conditions as independent. Typically, to compare the numer of spikes in two conditions we used a 3-way ANCOVA (injection amplitude and duration as ordinal variables; group id as a nominal variable) with interactions, and repeated measures (wieh cell id included as a nominal variable). For example, for a comparison of "Control" group to "Sound" group we used the following R code (see `dynamic_spikes_plot.R` for the full code):

```R
ds = subset(d,is.element(Group,c("Control","Sound")))
summary(aov(data=ds,Spikes~Shape*Amp + Shape*Group + Amp*Group + Cell*Shape*Amp))
```

We now included a brief description of our model in the "Results" section; improved the description in the "Methods" section; and added several referrals to the "Methods" section in the "Results" section, to hint that the reader may find full information there. In particular, in the sentence that reports the first F-value for this study, we now added the following commentary:

F(1,677)=30.4, p=5e-8, n cells=28, 29. Here and below, when F-values are reported, we use 4-way fixed effects analysis of variance with selected interactions, and cell id included as repeated measures factor, see "Methods" for a detailed description.

> 5.	Page 5 line 19 "This suggests that unlike the change in overall intrinsic excitability, the average temporal retuning was "adaptive" rather than "homeostatic", as neurons exposed to shorter stimuli (flashes) became more equipped to process shorter activation patterns." While it is appealing to give a name to this form of plasticity, the looming stimulus, which can be thought of as slower than the flash nonetheless caused a similar type of shift in response preference in the neurons, which runs counter the notion that there exists a meaningful dichotomy of adaptive vs. homeostatic change here.

> 6.	Figure 2B Amplitude is misspelled.

> 7.	The narrative logic on page 9 lines 11-22 is confusing. How does homeostatic plasticity, if it truly is at play here, "increase their ability to respond to unusual patterns of synaptic activation." If anything wouldn't homeostasis decrease dynamic range by pushing all cells toward a central set point?

> 8.	Page 12 the section on position within the tectum feels like it was just tacked onto the end of the paper. This should be integrated earlier in the paper perhaps to go with figure 1A.

@@

## Reviewer #2 (Rationale for Significance Rating for Authors (Required)):

> There is indeed some merit in the goals of the study (to see if adaptation changes with experience in addition to conductance/response gains), but it has so many technical problems at present that it would need to be radically different (perhaps with experiments and analysis re-done) to perhaps be of interest to the broad readership of J Neuroscience.

> The manuscript by Khakhalin and Busch describes the impact of experience on the intrinsic response properties of tectal neurons in the developing tadpole. The authors describe a surprising phenomenon: not only is the response gain of the conductance response altered by experience (in a homeostatic manner, if we imagine that the experience produces increased activity, which isn't shown), but the adaptation of the response is also impacted. The authors then go on to make some arguments, in my opinion unconvincing, that properties of the duration of the polysynaptic response to an optic chiasm shock are related to these intrinsic properties.

### 2018-12-03 main.tex
tex: stats descriptions

After conditioning, we excised the brain, obtained whole cell patch clamp recordings from neurons in the tectum (Figure 1A), and counted spikes produced in dynamic clamp, in response to simulated synaptic conductances of different durations and amplitudes (Figure 1B). We used conductances of 4 different durations (100, 200, 500, and 1000 ms), and 3 different amplitudes (peak conductances of 0.2, 0.5, and 1.0 nS), matching the range of synaptic currents observed in tectal circuits \textit{in vivo} (see "Methods"). Contrary to our expectations, in tadpoles that were exposed to instantaneous checkerboard inversions (flashes), tectal neurons on average became not more, but less spiky, and generated 0.4$\pm$0.4 spikes, across all types of dynamic clamp inputs, compared to 0.9$\pm$1.0 spikes in control (Figure 1D, left column; F(1,677)=30.4, p=5e-8, 4-way fixed effects linear model with interactions and repeated measures, see "Methods" for a detailed description of models; n cells=28, 29). Neurons exposed to looming transitions also spiked less than control neurons (0.6$\pm$0.4 spikes; F(1,629)=9.6; p=0.002; n=28, 25), but more than those exposed to instantaneous "flashes" (F(1,641)=3.7; p=4e-6; n=29, 25), which was likewise unexpected. In the tadpole tectum, looming stimuli are known to elicit stronger responses compared to instantaneous inversions \citep{khakhalin2014}, yet the suppression yielded by looming stimuli was weaker.

@@

To see whether prolonged sensory stimulation affected synaptic inputs received by tectal neurons, we recorded evoked excitatory postsynaptic currents in response to optic chiasm stimulation. We found that the amplitude of the early, monosynaptic component of evoked responses (the average current between 5 and 15 ms after the shock; Figure 3B) differed across experimental groups (Figure 3A; F(5,161)=3.2, p=0.009). Both Sync and Async multisensory groups had larger early synaptic currents than the Control group (Tukey p=0.03 and 0.04; Cohen d=0.46 and 0.74 on log-transformed data respectively). The amplitude of late synaptic currents–produced by recurrent network activation (15-145 ms after the stimulus)–did not differ across groups.

@@

To analyze the numbers of spikes observed in dynamic clamp experiments (Figure 1), we first averaged the number of generated spikes across 5 repetitions for each combination of conductance curve duration and amplitude, which resulted in 12 values per cell. Then we used sequential sum of squares analysis of variance with repeated measures. Both different condunctance curve amplitudes and durations were represented as ordinal values, and differences between experimental groups were assessed as interactions between these ordinal values and the factor variable encoding the experimental group. Cell id was included as a fixed factor interaction, as in repeated measures or blocking analysis of variance. To verify the validity of this approach, we also ran a maximal likelihood mixed-effects model with type III interaction terms, as implemented in R package “lmer”, with “lmerTest” extension to get access to Satterthwaite degrees of freedom and p-value estimations \citep{kuznetsova2017}. The results of both methods were numerically very similar.

For comparisons or summative values and individual cell parameters between experimental groups we report p-values of fixed effect sequential sum of squares linear model (ancova), in which rostral and medial coordinates of each cell within the tectum are included as covariates, and experimental group - as the main factor. All comparisons and correlations between cell parameters are performed on values corrected for cell position within the tectum. Position correction was based on a two-way linear regression model without interaction. For five variables that were distributed extremely non-normally, this correction for position was performed on transformed values (original values were transformed, linearly adjusted, and then transformed back): for early and late mean synaptic amplitudes we used a transformation $a'=log(1-a)$; for the variability of synaptic amplitudes $s'=log(1+s)$; for temporal tuning $y_t' = \sqrt{y_t}$ . Where appropriate, we performed the analysis with and without extreme outliers, and reported the difference. All analyses presented in the paper were also verified in mixed model analyses, with animal id included as a random factor.

>>>>

After conditioning, we excised the brain, obtained whole cell patch clamp recordings from neurons in the tectum (Figure 1A), and counted spikes produced in dynamic clamp, in response to simulated synaptic conductances of different durations and amplitudes (Figure 1B). We used conductances of 4 different durations (100, 200, 500, and 1000 ms), and 3 different amplitudes (peak conductances of 0.2, 0.5, and 1.0 nS), matching the range of synaptic currents observed in tectal circuits \textit{in vivo} (see "Methods"). Contrary to our expectations, in tadpoles that were exposed to instantaneous checkerboard inversions (flashes), tectal neurons on average became not more, but less spiky, and generated 0.4$\pm$0.4 spikes, across all types of dynamic clamp inputs, compared to 0.9$\pm$1.0 spikes in control (Figure 1D, left column; F(1,677)=30.4, p=5e-8, n cells=28, 29. Here and below, when F-values are reported, we use 4-way fixed effects analysis of variance with selected interactions, and cell id included as repeated measures factor, see "Methods" for a detailed description). Neurons exposed to looming transitions also spiked less than control neurons (0.6$\pm$0.4 spikes; F(1,629)=9.6; p=0.002; n=28, 25), but more than those exposed to instantaneous "flashes" (F(1,641)=3.7; p=4e-6; n=29, 25), which was likewise unexpected. In the tadpole tectum, looming stimuli are known to elicit stronger responses compared to instantaneous inversions \citep{khakhalin2014}, yet the suppression yielded by looming stimuli was weaker.

@@

To see whether prolonged sensory stimulation affected synaptic inputs received by tectal neurons, we recorded evoked excitatory postsynaptic currents in response to optic chiasm stimulation. We found that the amplitude of the early, monosynaptic component of evoked responses (the average current between 5 and 15 ms after the shock; Figure 3B) differed across experimental groups (Figure 3A; F(5,161)=3.2, p=0.009; see "Methods" for a description of the linear model we used). Both Sync and Async multisensory groups had larger early synaptic currents than the Control group (Tukey p=0.03 and 0.04; Cohen d=0.46 and 0.74 on log-transformed data respectively). The amplitude of late synaptic currents–produced by recurrent network activation (15-145 ms after the stimulus)–did not differ across groups.

@@

To analyze the numbers of spikes observed in dynamic clamp experiments (Figure 1), we first averaged the number of generated spikes across 5 protocol repetitions, for each cell, and for each combination of conductance curve duration and amplitude, which resulted in 12 values per cell. Then we used sequential sum of squares analysis of variance with repeated measures. Both different conductance curve amplitudes and durations were represented as ordinal values (1 to 3 for amplitudes, 1 to 4 for durations). Differences between experimental groups were assessed as interactions between these ordinal values and the factor variable encoding the experimental group, as we were interested in response shapes (reflected by interactions) rather than average values of spikiness (reflected by independent terms). Cell ids were included in the analysis as a fixed factor for repeated measures analysis of variance (also equivalent to analysis of variance with blocking). To verify the validity of this approach, we also ran a maximal likelihood mixed-effects model with type III interaction terms, as implemented in R package “lmer”, with “lmerTest” extension to get access to Satterthwaite degrees of freedom and p-value estimations \citep{kuznetsova2017}. The results of both methods were numerically very similar.

For the comparison or summative descriptions of tuning, and other electrophysiological cell parameters between experimental groups, we report p-values of fixed effect sequential sum of squares linear model (ancova), in which rostral and medial coordinates of each cell within the tectum are included as covariates, and experimental group is used as the main factor. All comparisons and correlations between cell parameters are performed on values corrected for cell position within the tectum. Position correction was based on a two-way linear regression model without interaction. For five variables that were distributed extremely non-normally, this correction for position was performed on transformed values (original values were transformed to normally distributed proxy values, linearly adjusted, and then transformed back): for early and late mean synaptic amplitudes we used a transformation $a'=log(1-a)$; for the variability of synaptic amplitudes $s'=log(1+s)$; for temporal tuning $y_t' = \sqrt{y_t}$ . Where appropriate, we performed the analysis with and without extreme outliers, and reported the difference. All analyses presented in the paper were also verified in mixed model analyses, with animal id included as a random factor; the results of these mixed models analyses were similar to that of a fixed model, and are not reported.
